---
title: "Including Stata (alternative)"
format: html
engine: knitr
---

```{r}
#| include: false
library(Statamarkdown)
```

As mentioned before, the `nbstata` approach to incorporating Stata code and output in Quarto documents does not permit parallel execution of other languages. In this section, I will demonstrate a solution that uses the R-package `Statamarkdown` in place of the Python-package `nbstata`. 

An alternative solution is offered by Dr Tom Palmer in the [R-bloggers post](https://www.r-bloggers.com/2025/02/creating-effectively-multi-engine-quarto-documents-using-quartos-embed-shortcode/). It exploits the fact that you can compile .qmd files using different engines/kernels and then embed the output into the same file. My reason for skipping this solution is that you would need to write a separate file for each language **and codeblock**.

Instead, I am following a solution alluded to in the above blog using `Statamarkdown`: a R-package designed to execute Stata code within R-markdown documents. 

## Setup

You will need to install the following package in R. 

```{r}
#| eval: false
install.packages("Statamarkdown")
```

Then, in your .qmd file you will need to add the following: 
````
```{{r}}
#| include: false
library(Statamarkdown)
```
````

This will load the package needed to execute the .qmd file. 

## Implementation

Once, the package is loaded the implementation is relatively straight foward. As before, you can setup tabs using tabset. 

````
:::{.panel-tabset group="languages"}

## R

```{{r}}
print("Hello world!")
```

## Stata

```{{stata}}
dis "Hello world!"
```
:::
````

The output will look as follows:

:::{.panel-tabset group="languages"}

## R

```{r}
print("Hello world!")
```

## Stata

```{stata}
dis "Hello world!"
```
:::

## Limitation

A key limitation of `Statamarkdown` is that it each codeblock is executed as a separate .do file. This means that memory is lost between codeblocks. 

:::{.panel-tabset group="languages"}

## R

```{r}
test <- "Am I still defined in R?"
test
```

## Stata

```{stata}
global test "Am I still defined in Stata?"
display "$test"
```
:::

Now, in a separate codeblock we try to display the same objects:

:::{.panel-tabset group="languages"}

## R

```{r}
test
```

## Stata

```{stata}
dis "$test No"
```
:::

Note, this problem is specific to Stata and not R. 

## Solution

You need to add `collectcode=TRUE` to ensure that the Stata codes are not run as separate files. 

````
:::{.panel-tabset group="languages"}

## R

```{{r}}
print("Hello world!")
```

## Stata

```{{stata, collectcode=TRUE}}
dis "Hello world!"
```
:::
````
Alternatively, you can set this globally at the top of the .qmd file by adding the following code in R:

````
```{{r}}
#| include: false
knitr::opts_chunk$set(collectcode = TRUE)
```
````
In the next section I will implement this change. Switching `collectcode` on part way through a file causes trouble in the rendering. 

:::{.callout-note title="Statamarkdown vs nbstata"}
Statamarkdown's implementation is not as neat as nbstata. The key limitation is that each codeblock appears to be executed as a separate (temporary) do-file, which means local macros are lost across codeblocks, even when `collectcode=TRUE` (see next page for example).

That said, Statamarkdown works with tabset, which is a very nice feature for resource design. 
:::